---
title: "Project2 - EDA"
author: "Qing Gao"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggmap)
library(devtools)
library(choroplethrZip)

```


```{r, echo = FALSE}
# Load data
df <- read.csv('./data/DOHMH_New_York_City_Restaurant_Inspection_Results.csv')
```

```{r, echo = FALSE}
# Data cleaning: convert names into lower cases
df.clean <- df %>% 
  rename_all(tolower) %>%
  rename(cuisine = cuisine.description) %>%
  mutate(cuisine = str_replace(cuisine, " \\(.*\\)", "")) %>%
  mutate(region = as.character(zipcode)) %>%
  filter(region > 0 & region != 'N/A')


```

## EDA
- Cuisine Type
- Grade & Score
- Location

### Cuisine Type
```{r, echo = FALSE}
# Count and arrange the frequency of the cuisine type
cuisine <- df.clean %>%
  group_by(cuisine) %>%
  count() %>%
  tally(n, sort = T)

# Combine all rows into one row named 'Other' except for the first 10 rows,
cuisine_10 <- cuisine %>%
  group_by(cuisine = factor(c(cuisine[1:10], rep("Other", n() - 10)),
                            levels = c(cuisine[1:10], "Other"))) %>%
  tally(n)

# Compute percentages & cumulative percentages
cuisine_10 <- cuisine_10 %>%
  mutate(frac = n / sum(n), ymax = cumsum(frac), ymin = c(0, head(ymax, n=-1)),
         label = paste0(cuisine, "\n", n),
         labelPosition = (ymax + ymin) / 2)

# Data visualization - donut chart: cuisine distribution
ggplot(cuisine_10, aes(ymax=ymax, ymin=ymin, xmax=11, xmin=10, fill=cuisine)) +
  geom_rect() +
  geom_label(x=11.4, aes(y=labelPosition, label=label), size=2.2) +
  scale_fill_brewer(palette = "Set3") +
  scale_color_brewer(palette = "Set3") +
  coord_polar(theta="y") +
  xlim(c(7, 11)) +
  labs(x = '', y = '', title = 'Distribution of Cuisine Type') +
  theme_light() +
  theme_void()  
```

The most popular types of restaurant in our datasets are American, Chinese, Cafe, Latin and Pizza.


### Grade & Score
```{r, echo = FALSE}
# Summarize the grade by cuisine type
grade <- df.clean %>%
  select(cuisine, grade) %>%
  filter((grade %in% c('A', 'B', 'C', 'P', 'Z')) &
          cuisine %in% cuisine_10$cuisine[1:10]) %>%
  group_by(cuisine, grade) %>%
  count() %>%
  group_by(cuisine) %>%
  mutate(frac = n / sum(n))
  

# Data visualization - heat map
ggplot(grade, aes(grade, cuisine, fill = frac)) +
  geom_tile(alpha = 0.8) +
  scale_fill_gradient(low  = "white", high="dark blue") +
  labs(x = 'Grade', y = 'Cuisine',
       title = 'Performance of Different Types of Restaurants on Grade')

```

From the heatmap, it is not obvious to show the performance of each type of restaurants on the inspection grade. Latin, and Caribbean restaurants have less A grade comparing to other restaurants. 


### Location
```{r, echo = FALSE}
# Count the restaurants by zipcode
region <- df.clean %>%
  group_by(region) %>%
  summarise(value = n())

# Draw the map of the restaurant distribution
nyc = c(36005, 36047, 36061, 36081, 36085)
zip_choropleth(region,
    title       = "New York Restaurant",
    legend      = "Number of Restaurant",
    county_zoom = nyc)

```



